### 2022-01-11 paper/paper.tex
methodology: add initial content

@@

Capturing of the video to be analyzed by the program would be limited to a
single 1080p webcam that is capturing in 60 frames per second. The camera will
be pointed downwards facing the keyboard and the hand. This means that the
vertical angle of the wrist may not be accurate.

@@

\subsubsection{MediaPipe}

@@

\citeauthor{dorf2001} presents a use case for finger tracking in augmented
environments. In the paper, interaction in a virtual environment through the use
of gestures. The tracking system uses an optical marker-based motion capture
system where the user wears a glove with retroreflective markers.

\citeauthor{chiang2014} used a Kinect, a 3D sensing device by Microsoft that
uses depth data, to track fingers to play virtual instruments. Virtual Pianos
and Guitars were created and played with reliable and stable tracking.

\citeauthor{yousaf2014} created a virtual keyboard that operates using finger
tracking. The tracking uses the movement of the finger joints as the basis for
selecting which key to press. A camera captures the movement, and the resulting
video stream is used for hand region detection and finger joint localization.
Using probabilistic regional density-based kernel tracking, finger joint
trajectories are gathered. Feature vectors are then interpreted from the
trajectories. These feature vectors are used in logic-based techniques and
Dynamic Bayesian Network for classification, detection, and recognition of
keystrokes.

@@

>>>>

% Misc
\usepackage{gensymb}

%Flowchart
\usepackage{tikz}
\usetikzlibrary{shapes,positioning,calc}

@@

Capturing of the video to be analyzed by the program would be limited to a
single 480p webcam that is capturing in 24 frames per second. The camera will
be pointed downwards facing the keyboard and the hand. This means that the
vertical angle of the wrist may not be accurate.

@@

\subsubsection{MediaPipe}
\label{section:rrl-mediapipe}

@@

\cite{dorf2001} presents a use case for finger tracking in augmented
environments. In the paper, interaction in a virtual environment through the use
of gestures. The tracking system uses an optical marker-based motion capture
system where the user wears a glove with retroreflective markers.

\cite{chiang2014} used a Kinect, a 3D sensing device by Microsoft that
uses depth data, to track fingers to play virtual instruments. Virtual Pianos
and Guitars were created and played with reliable and stable tracking.

\cite{yousaf2014} created a virtual keyboard that operates using finger
tracking. The tracking uses the movement of the finger joints as the basis for
selecting which key to press. A camera captures the movement, and the resulting
video stream is used for hand region detection and finger joint localization.
Using probabilistic regional density-based kernel tracking, finger joint
trajectories are gathered. Feature vectors are then interpreted from the
trajectories. These feature vectors are used in logic-based techniques and
Dynamic Bayesian Network for classification, detection, and recognition of
keystrokes.

@@

\chapter{Methodology}

\section{Setup}
\label{section:metho-setup}

The experimental setup and configuration will be composed of three elements: the
camera, the keyboard, and the environment.

\subsection{Camera}
The camera setup will use a single monocular camera positioned in a top-down
view. The camera should capture the entirety of the keyboard and the movement of
the ten fingers. Figure~\ref{fig:metho-keeb} shows what the camera will capture
once placed in its correct position.

The specific camera to be used will be a Logitech C270.
Figure~\ref{fig:metho-cam} is a picture of this camera. This is a 0.9 mega pixel
webcam with a CMOS sensor. This camera is capable of capturing color videos in
720p/30fps and it has a diagonal field of view of 55\degree. This camera has a
universal mounting clip that will reduce the difficulty in mounting the camera
\parencite{logitech}.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.4\textwidth]{webcam.png}
	\caption{Logitech C270. Reprinted from \fullcite{logitech}}
	\label{fig:metho-cam}
	\centering
\end{figure}

\subsection{Keyboard}
\label{section:metho-keeb}
The keyboard will be a 60\% keyboard as shown in Figure~\ref{fig:metho-keeb}.
This will limit the necessary mapping for the algorithm to the alphanumeric
portion of the keyboard. This keyboard choice will also lessen the area that the
camera will need to capture, as this keyboard type is considerably smaller
compared to a full-size keyboard. The keyboard will have key caps and a case in
a light color that will contrast the dark surface that it will be placed on.
This is to improve initial keyboard detection. In addition, the keyboard layout
will also be ANSI, due to the availability and widespread adoption of the layout
in the Philippines.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{actual-keeb.png}
	\caption{Keyboard, camera angle, and placement to be used for testing.}
	\label{fig:metho-keeb}
	\centering
\end{figure}

\subsection{Environment}
The environment the setup would be placed in would be a well lit environment
with a light source beside the camera. This light source will evenly light the
keyboard and the fingers used for typing. The light source to be used will be a
common LED desk lamp rated at 9 Watts with 700 lumens. This light will be white
with a color temperature of 6500k.

In addition, the surface where the keyboard is to be placed on will be solid
green without any variation of color. This is to improve initial keyboard
detection.

\section{Algorithm}
\label{section:metho-algo}

\subsection{Computer Vision based Keyboard Detection, and Mapping}
\label{section:metho-algo-keyboard}

A computer vision algorithm will be created as a starting point for detecting,
and mapping the keyboard within a video. The flowchart of the algorithm is shown
in Figure~\ref{fig:metho-key-algo}.

\begin{figure}[H]
	\centering
	\begin{tikzpicture}[font=\small,thick]
		\node[draw,
			rounded rectangle,
			minimum width=2.5cm,
			minimum height=1cm] (start) {START};

		\node[draw,
			trapezium,
			trapezium left angle = 65,
			trapezium right angle = 115,
			trapezium stretches,
			below=of start,
			minimum width=3.5cm,
			minimum height=1cm
		] (input) {Input Frame};

		\node[draw,
			below=of input,
			minimum width=3.5cm,
			minimum height=1cm] (bg-rem) {Background Removal};

		\node[draw,
			below=of bg-rem,
			minimum width=3.5cm,
			minimum height=1cm] (gray) {Convert to Grayscale};

		\node[draw,
			below=of gray,
			minimum width=3.5cm,
			minimum height=1cm] (sobel) {Edge Detection using Sobel Filter};

		\node[draw,
			below=of sobel,
			minimum width=3.5cm,
			minimum height=1cm] (contours) {Find Contours (Simple)};

		\node[draw,
			below=of contours,
			minimum width=3.5cm,
			minimum height=1cm] (map) {Overlay Virtual Keyboard Map using Contour Points};

		\node[draw,
			trapezium,
			trapezium left angle = 65,
			trapezium right angle = 115,
			trapezium stretches,
			below=of map,
			minimum width=3.5cm,
			minimum height=1cm
		] (output) {Output Frame};

		\node[draw,
			below=of output,
			rounded rectangle,
			minimum width=2.5cm,
			minimum height=1cm] (end) {END};

		\draw[-latex] (start) edge (input)
		(input) edge (bg-rem)
		(bg-rem) edge (gray)
		(gray) edge (sobel)
		(sobel) edge (contours)
		(contours) edge (map)
		(map) edge (output)
		(output) edge (end);

	\end{tikzpicture}
	\caption{Flowchart of Keyboard Detection and Mapping Algorithm}
	\label{fig:metho-key-algo}
\end{figure}

\subsubsection{Background Removal}
The background of the frame will be removed by using colors as the main metric
for determining the foreground and the background. This is made easier by the
setup as the setup will have a green background, with a light colored keyboard
as a foreground.

\subsubsection{Convert to Grayscale}
The frame will then be converted to grayscale. This is done because future steps
of the algorithm does not require color values and this improves the performance
of these algorithms.

\subsubsection{Edge Detection using Sobel Filter}
A Sobel Filter is a edge detection filter that uses a 3$times$3 kernel that is
convolved twice. Once horizontally, and another vertically to produce a
grayscale image of the outlines within the frame.

\subsubsection{Find Contours (Simple)}
Contours are curves joining all continuous points that have the same color or
intensity \parencite{opencv-contours}. In this algorithm, an OpenCV function is
to be used to find the contours of the outlines of the object found using the
previous step. This OpenCV function implements the algorithm of \cite{contours}.
The expected result is four contour points which are positioned on the four
edges of the keyboard.

\subsubsection{Overlay Virtual Keyboard Map using Contour Points}
The virtual keyboard map is a rectangular image that has individual region of
interest (ROI) for each key in a 60\% \ac{ansi} keyboard. Each key has a
corresponding color assigned to it and this color fills the region the key is
assigned to.

The virtual map is then stretched over the object, with the four contour points
as the four edges of the virtual keyboard map. The image generated is then
returned as the final output of the graph.

\subsection{Computer Vision based Finger Detection, and Tracking}
\label{section:metho-algo-finger}

A computer vision algorithm will be chosen as a starting point for detecting,
and tracking fingers within a video.

The first component this algorithm will have is the ability to detect fingers
within a video. This would involve capturing exactly where the finger is in the
current frame. This data would then be used for the next component of the
algorithm: tracking. This component would track the movement of the finger as it
performs a typing motion. This algorithm should have the ability to separate the
different fingers and track the movement of the ten fingers individually.

The specific algorithm used for finger detection and tracking will be the
solution offered by MediaPipe. More details about the specific algoritihm can be
found in Section~\ref{section:rrl-mediapipe}.

\subsection{Integration for finger-key identification and mapping}

The two previously chosen algorithms will be combined to accomplish finger-key
identification and mapping. Finger-key identification refers to identifying
which finger is used to press which key. As an example, the key Q was pressed
with the pinky finger of the left hand.

The integration of the algorithm is a two step process. The first step is to get
the mapping of the keyboard using the algorithm shown in Section~\ref{section:metho-algo-keyboard}.

The second step is a continous loop where the mapping of the keyboard is used in
conjunction with the finger tracking algorithm chosen in
Section~\ref{section:metho-algo-finger}. The flowchart of the algorithm is shown
in Figure~\ref{metho-algo-integration}

\begin{figure}[H]
	\centering
	\begin{tikzpicture}[font=\small,thick]
		\node[draw,
			rounded rectangle,
			minimum width=2.5cm,
			minimum height=1cm] (start) {START};

		\node[draw,
			trapezium,
			trapezium left angle = 65,
			trapezium right angle = 115,
			trapezium stretches,
			below left=of start,
			minimum width=3.5cm,
			minimum height=1cm
		] (map) {Virtual Keyboard Map};

		\node[draw,
			trapezium,
			trapezium left angle = 65,
			trapezium right angle = 115,
			trapezium stretches,
			below right=of start,
			minimum width=3.5cm,
			minimum height=1cm
		] (track) {Finger Tracking Frame};

		\coordinate (CENTER) at ($(map)!0.5!(track)$);

		\node[draw,
			diamond,
			aspect=2,
			below = 3cm of CENTER,
			minimum width=2.5cm,
			inner sep=0] (if-packet) {Keypress Packet Received};

		\node[draw,
			below left=of if-packet,
			minimum width=3.5cm,
			minimum height=1cm] (find-roi) {Find ROI of the pressed key in Virtual Keyboard Map};

		\node[draw,
			below=of find-roi,
			minimum width=3.5cm,
			minimum height=1cm] (find-finger) {Find which fingertip is on that ROI};

		\node[draw,
			trapezium,
			trapezium left angle = 65,
			trapezium right angle = 115,
			trapezium stretches,
			below=of find-finger,
			minimum width=3.5cm,
			minimum height=1cm
		] (return) {Return which fingertip};

		\node[draw,
			trapezium,
			trapezium left angle = 65,
			trapezium right angle = 115,
			trapezium stretches,
			below = 11cm of CENTER,
			minimum width=3.5cm,
			minimum height=1cm
		] (output) {Output Frame};

		\node[draw,
			below=of output,
			rounded rectangle,
			minimum width=2.5cm,
			minimum height=1cm] (end) {END};

		\draw[-latex] (start) edge (map)
		(start) edge (track)
		(track) edge (if-packet)
		(map) edge (if-packet)
		(find-roi) edge (find-finger)
		(find-finger) edge (return)
		(return) edge (output)
		(output) edge (end);

		\draw[-latex] (if-packet) -| (find-roi)
		node[pos=0.25,fill=white,inner sep=0]{true};

		\draw[-latex] (if-packet) edge node[pos=0.4,fill=white,inner sep=2pt]{No}(output);

	\end{tikzpicture}
	\caption{Flowchart of the overall flow}
	\label{fig:metho-algo-integration}
\end{figure}

\subsubsection{Find ROI of the pressend key in Virtual Keyboard Map}
Since there is a known color-key mapping based on Virtual Keyboard Map, the
algorithm will find the correct color to search for based on the received keypress
packet. This color is isolated from the Virtual Keyboard Map, and the region of
interest and its coordinates are obtained.

\subsubsection{Find which fingertip is on that ROI}
The Finger Tracking Frame contains the pixel positions of each landmark of the
hand. For this step, the algorithm will find the landmark which is positioned
within the region of interest from the previous step.

\subsubsection{Return which fingertip}
The finger containing the landmark will be returned. This finger will then be
compared with the expected finger to be used to press the key to determine if it
is correct or not.

\section{Trainer}

\subsection{Create typing test sequences}
Typing test sequences are strings that will be used in testing the user in their
ability to type. The test sequences that will be used in the trainer will come
from text found in the public domain obtained from Project Gutenberg and the
Internet Archive. Sentences from these text will be isolated and used as the
test sequences. There will be a total of 20 test sequences that will be
gathered. An example test sequence is ``What of it, if some old hunks of a
sea-captain orders me to get a broom and sweep down the decks?'' from Moby Dick
by \cite{moby-dick}. The 19 other test sequences can be found in the appendix.

\subsection{UX/UI design and Front-end development for the touch typing trainer}
The touch typing trainer will have its UX/UI design created in Figma. This
design will be based on other type tests such as Monkeytype
\parencite{bartnik2021}, and Keybr \parencite{keybr}, with additional components
for showing real-time finger-key identification and mapping.
Figure~\ref{fig:ui-test} and \ref{fig:ui-stat} illustrates the two main pages of
the design.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{ui-test.png}
	\caption{Test screen of the front-end design}
	\label{fig:ui-test}
	\centering
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{ui-stats.png}
	\caption{Statistics screen of the front-end design}
	\label{fig:ui-stat}
	\centering
\end{figure}

The front-end will be web based, and the framework the front-end will be
developed in will be Svelte, due to its community resources, performance, and
the researcher's familiarity with the framework.

\subsection{Back-end development of a touch typing trainer using the program created
	in \ref{section:metho-algo} and the experimental setup identified in
	\ref{section:metho-setup}}
\subsection{Testing of finger-key identification and mapping accuracy}
\section{Metric}
%% 4. Finger Placement Keyboard Typing Metric Development
%%     1. Determine if the metric is for individual key presses or for the whole typing sequence or both.
%%     2. Evaluate which variables need to be taken into account for the metric depending on (5.a):
%%         1. Actual Finger Used
%%         2. Correct Finger to be Used
%%         3. Specific key pressed
%%         4. Number of keys pressed with the correct fingers
%%         5. Number of keys pressed with the wrong fingers
